---
title: "EyetrackingDataCleanUp"
author: "RF"
date: "2/1/2018"
output: html_document
---

# Here we load libraries and data (from 2018)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readr)
library(groupdata2)

setwd('/home/tnoncs/eyetracking')
Samples=read_delim('eyetracking_2018_samples.txt',delim='\t')
Fixations=read_delim('eyetracking_2018_fixations.txt',delim='\t')
Saccades=read_delim('eyetracking_2018_saccades.txt',delim='\t')
```

# Here we clean up the data
- We only select useful columns
- We rename the columns in more concise and meaningful ways
- We add the task (Visual Search vs. Social Engagement)
- We add time as counted from the start of each trial
- We check whether the trials are of the length we expect (6s for videos, 10s/20s/40s for images)

```{r}
AddTask <- function(df){
  for (s in unique(df$ParticipantID)){
    if (max(df$Trial[df$ParticipantID==s])==8){
      df$Task[df$ParticipantID==s]='SocialEngagement'
    } else {
      df$Task[df$ParticipantID==s]='VisualSearch'
    }
  }
  df$ParticipantGender[grepl("f|F",df$ParticipantID)]='female'
  df$ParticipantGender[grepl("m|M",df$ParticipantID)]='male'
  
  df$SearchOrder[grepl("f1|F1",df$ParticipantID)]=1
  df$SearchOrder[grepl("m1|M1",df$ParticipantID)]=1
  df$SearchOrder[grepl("f2|F2",df$ParticipantID)]=2
  df$SearchOrder[grepl("m2|M2",df$ParticipantID)]=2
  
  return(df)
}
library(dplyr)
Fixations = Fixations %>% 
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    Fixation=CURRENT_FIX_INDEX,
    Duration=CURRENT_FIX_DURATION,
    PupilSize=CURRENT_FIX_PUPIL,
    PositionX=CURRENT_FIX_X,
    PositionY=CURRENT_FIX_Y,
    StartTime=CURRENT_FIX_START,
    EndTime=CURRENT_FIX_END)


Fixations=AddTask(Fixations)


Saccades = Saccades %>%
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    Saccade=CURRENT_SAC_INDEX,
    Direction=CURRENT_SAC_DIRECTION,
    Amplitude=CURRENT_SAC_AMPLITUDE,
    Duration=CURRENT_SAC_DURATION,
    PeakVelocity=CURRENT_SAC_PEAK_VELOCITY,
    MeanVelocity=CURRENT_SAC_AVG_VELOCITY,
    Blink=CURRENT_SAC_CONTAINS_BLINK,
    StartTime=CURRENT_SAC_START_TIME,
    EndTime=CURRENT_SAC_END_TIME,
    StartX=CURRENT_SAC_START_X,
    StartY=CURRENT_SAC_START_Y,
    EndX=CURRENT_SAC_END_X,
    EndY=CURRENT_SAC_END_Y
  )
Saccades$Amplitude=as.numeric(Saccades$Amplitude)
Saccades$MeanVelocity=as.numeric(Saccades$MeanVelocity)
Saccades$EndX=as.numeric(Saccades$EndX)
Saccades$EndY=as.numeric(Saccades$EndY)

Saccades=AddTask(Saccades)

Samples = Samples %>%
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    TimeStamp=TIMESTAMP,
    PositionX=RIGHT_GAZE_X,
    PositionY=RIGHT_GAZE_Y,
    PupilSize=RIGHT_PUPIL_SIZE,
    Saccade=RIGHT_IN_SACCADE,
    Blink=RIGHT_IN_BLINK,
    MeanVelocityX=RIGHT_VELOCITY_X,
    MeanVelocityY=RIGHT_VELOCITY_Y,
    MeanAccellerationX=RIGHT_ACCELLERATION_X,
    MeanAccellerationY=RIGHT_ACCELLERATION_Y
  )

Samples$MeanVelocityY=as.numeric(Samples$MeanVelocityY)
Samples$MeanVelocityX=as.numeric(Samples$MeanVelocityX)
Samples$MeanAccellerationY=as.numeric(Samples$MeanAccellerationY)
Samples$MeanAccellerationX=as.numeric(Samples$MeanAccellerationX)

Samples=AddTask(Samples)

# Add time from trial start

Samples <- Samples %>%
  group_by(ParticipantID,Trial) %>%
  mutate(TrialTime = TimeStamp - min(TimeStamp))

## Double check trial timing. It's all in place!
X <- Samples %>%
   group_by(ParticipantID,Trial) %>%
   dplyr::summarize(LengthSeconds = max(TrialTime)/1000)
print(X)
```


## Import fixation number and saccade number from processed data to samples
Since the samples do not exclude artifacts and do not contain fixation and saccade number, we import this data from Fixations and Saccades datasets

```{r}

for (t in seq(nrow(Fixations))){
  StartTime=Fixations$StartTime[t]
  EndTime=Fixations$EndTime[t]
  ID=Fixations$ParticipantID[t]
  Trial=Fixations$Trial[t]
  FixN=Fixations$Fixation[t]
  Samples$Fixation[Samples$TrialTime >= StartTime  & Samples$TrialTime <= EndTime & Samples$ParticipantID==ID & Samples$Trial==Trial]=FixN
}


for (t in seq(nrow(Saccades))){
  StartTime=Saccades$StartTime[t]
  EndTime=Saccades$EndTime[t]
  ID=Saccades$ParticipantID[t]
  Trial=Saccades$Trial[t]
  SacN=Saccades$Fixation[t]
  Samples$SaccadeN[Samples$TrialTime >= StartTime  & Samples$TrialTime <= EndTime & Samples$ParticipantID==ID & Samples$Trial==Trial]=SacN
}

write_csv(Saccades,"SaccadesV1.csv")
write_csv(Fixations,"FixationsV1.csv")
write_csv(Samples,"SamplesV1.csv")

```

## Merge with data about experimental conditions
We merge the data with the experimental logs.

```{r}
Fixations=read_csv('FixationsV1.csv')
Saccades=read_csv('SaccadesV1.csv')
Samples=read_csv('SamplesV1.csv')

## THEN GO TO THE LOG FILES AND MATCH the data from the logfiles to the samples/fixations/saccades (to have conditions, and so forth)
## ADD GROUP PARTICIPANT AND GENDER TO THE TABLES

# function to get out the info from the video file names
get_info= function(name){
  nev=name [[1]]  #rows in the videos column
  ActorGender= substring(nev, 1, 1)
  Directionality= substring(nev, 9, 11)
  Ostension= substring(nev, 13, 14)
  save= cbind(ActorGender, Directionality, Ostension) 
  return(save)
}

#read in everything
file_list = list.files(path= "PupilsLogs", full.names=T)
all_files=lapply(file_list, FUN=read.csv, header=TRUE)

# works
z=list(1,2,3,4,5,6)
x=1
videos=data.frame()

for (y in z) {
  y= z[[x]]
  ParticipantID= all_files[[y]]$subject
  ID= as.data.frame(ParticipantID) #save the ID to be able to merge
  Trial= all_files[[y]]$X
  tri= as.data.frame(Trial)
  tri= tri+1 #save trial number - +1, mert eggyel kezdodne
  
  file_name= all_files[[y]]$video
  fn= as.data.frame(file_name) #it's a dataframe with one column with the video file names
  info= get_info(fn) 
  
  final= cbind(ID, tri, info)
  videos= rbind(videos, final)
  x=x+1
  }

stim1=read.table("ForagingOrder1.txt", col.names =  "Stimulus")
stim2=read.table("ForagingOrder2.txt", col.names =  "Stimulus")

searchtype1= data.frame(c(rep("Count", times=5), rep("Search", times=5)))
colnames(searchtype1) = "SearchType"

searchtype2= data.frame(c(rep("Search", times=5), rep("Count", times=5)))
colnames(searchtype2) = "SearchType"

type1= cbind(stim1, searchtype1)
type2= cbind(stim2, searchtype2)

type1$SearchOrder=1
type2$SearchOrder=2
type1$Trial=1:10
type2$Trial=1:10

type= rbind(type1, type2)

Fixations2= merge(Fixations, type, all=T)
Saccades2= merge(Saccades, type, all=T)
Samples2= merge(Samples, type, all=T)

Fixations3= merge(Fixations2, videos, all = T)
Saccades3= merge(Saccades2, videos, all=T)
Samples3= merge(Samples2, videos, all=T)

write.csv(Saccades3,"SaccadesV2.csv")
write.csv(Fixations3,"FixationsV2.csv")
write.csv(Samples3,"SamplesV2.csv")

```

```{r}

Order1<- read_tsv('ForagingOrder.txt',col_names = F)
Order1$Trial=1:10
Order1$SearchOrder=1
Order1$SearchType=c("Search","Search","Search","Search","Search","Count","Count","Count","Count","Count")
Order2<- read_tsv('ForagingOrder.txt',col_names = F)
Order2$Trial=1:10
Order2$SearchOrder=2
Order2$SearchType=c("Count","Count","Count","Count","Count","Search","Search","Search","Search","Search")
Order=rbind(Order1,Order2)
Order=Order %>% rename(Stimulus=X1)

Saccades=merge(Saccades,Order,all=T)
Fixations=merge(Fixations,Order,all=T)
Samples=merge(Samples,Order,all=T)

library(stringr)
setwd('home/tnoncs/eyetracking/PupilsLogs')
fileList<-list.files(".",pattern="csv")
logs=NULL
for (f in fileList){
  d<-read_csv(f)
  d=d %>% rename(Trial=X1,ParticipantID=subject)
  d$Trial=d$Trial+1
  for (n in seq(nrow(d))){
    x<-str_split(d$video[n],"_")[[1]]
    d$ActorGender[n]=x[1]
    d$Directionality[n]=x[4]
    d$Ostension[n]=x[5]
  }
  if (exists('logs')){logs=rbind(logs,d)} else {logs=d}
}

Saccades=merge(Saccades,logs,all=T)
Saccades$Stimulus[Saccades$Task=='SocialEngagement']=Saccades$video[Saccades$Task=='SocialEngagement']
Saccades$video=NULL
Fixations=merge(Fixations,logs,all=T)
Fixations=Fixations[!is.na(Fixations$Task),]
Fixations$Stimulus[Fixations$Task=='SocialEngagement']=Fixations$video[Fixations$Task=='SocialEngagement']
Fixations$video=NULL
Samples=merge(Samples,logs,all=T)
Samples$Stimulus[Samples$Task=='SocialEngagement']=Samples$video[Samples$Task=='SocialEngagement']
Samples$video=NULL
setwd('home/tnoncs/eyetracking')
write_csv(Saccades,"SaccadesV2.csv")
write_csv(Fixations,"FixationsV2.csv")
write_csv(Samples,"SamplesV2.csv")
```

# We now integrate the datasets with all the available data from 2017

```{r}

Fixations=read_csv('FixationsV2.csv')
Saccades=read_csv('SaccadesV2.csv')
Samples=read_csv('SamplesV2.csv')

ForagingFixations <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/FORAGING_FixationsClean.csv")
ForagingFixations$Task='VisualSearch'
ForagingSaccades <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/FORAGING_SaccadesClean.csv")
ForagingSaccades$Task='VisualSearch'
PupilsFixations <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/PUPIL_FixationsClean.csv")
PupilsFixations$Task='SocialEngagement'
PupilsSamples <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/PUPIL_SamplesClean.csv")
PupilsSamples$Task='SocialEngagement'


ForagingFixations <- ForagingFixations %>% rename(
  SearchType=Condition,
  SearchOrder=Order,
  Duration=Duration,
  StartTime=CURRENT_FIX_START,
  EndTime=CURRENT_FIX_END,
  PositionX=CURRENT_FIX_X,
  PositionY=CURRENT_FIX_Y
)

ForagingFixations[,c('PupilSize','ActorGender','Directionality','Ostension','ParticipantGender')]=NA

PupilsFixations <- PupilsFixations %>% rename(
  ParticipantID=participant_id,
  Trial=trial,
  Fixation=current_fix_index,
  Duration=current_fix_duration,
  StartTime=current_fix_start,
  EndTime=current_fix_end,
  PositionX=current_fix_x,
  PositionY=current_fix_y,
  Task=Task,
  PupilSize=current_fix_pupil,
  Stimulus=video,
  ActorGender=gender,
  Directionality=directed,
  Ostension=ostension
)

PupilsFixations[,c('Condition','SearchType','SearchOrder','ParticipantGender')]=NA

Fixations=Fixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]

ForagingFixations=ForagingFixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]

PupilsFixations=PupilsFixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]


Fixations2017=rbind(ForagingFixations,PupilsFixations)
Fixations2017$Year='2017'
Fixations$Year='2018'
Fixations=rbind(Fixations,Fixations2017)

ForagingSaccades <- ForagingSaccades %>% rename(
  SearchType=Condition,
  SearchOrder=Order,
  Direction=CURRENT_SAC_DIRECTION,
  Amplitude=CURRENT_SAC_AMPLITUDE,
  PeakVelocity=CURRENT_SAC_PEAK_VELOCITY,
  MeanVelocity=CURRENT_SAC_AVG_VELOCITY,
  Blink=CURRENT_SAC_CONTAINS_BLINK,
  StartTime=CURRENT_SAC_START_TIME,
  EndTime=CURRENT_SAC_END_TIME,
  StartX=CURRENT_SAC_START_X,
  StartY=CURRENT_SAC_START_Y,
  EndX=CURRENT_SAC_END_X,
  EndY=CURRENT_SAC_END_Y
)

ForagingSaccades[,c("ParticipantGender","Stimulus","ActorGender","Directionality","Ostension" )]=NA
ForagingSaccades$Year='2017'
Saccades=Saccades[,c("ParticipantID","Trial","Saccade","Direction","Amplitude","PeakVelocity","MeanVelocity","StartTime","EndTime","Duration", "StartX", "StartY","EndX","EndY","Blink","SearchType","SearchOrder","Task","ParticipantGender","Stimulus", "ActorGender" , "Directionality","Ostension")]
Saccades$Year='2018'  
Saccades=rbind(Saccades,ForagingSaccades)

PupilsSamples <- PupilsSamples %>% rename(
  ParticipantID=participant_id,
  Trial=trial,
  PupilSize=right_pupil_size,
  Fixation=fix_nr,
  Saccade=right_in_saccade,
  Blink=right_in_blink,
  ActorGender=gender,
  Directionality=directed,
  Ostension=ostension,
  Stimulus=video,
  TrialTime=task_time,
  Task=Task
)

PupilsSamples[,c("SearchOrder","TimeStamp","PositionX","PositionY","MeanVelocityX","MeanVelocityY","MeanAccellerationX","MeanAccellerationY","ParticipantGender","SearchType")]=NA
PupilsSamples$fixation=NULL

Samples=Samples[,colnames(PupilsSamples)] 
PupilsSamples$Year='2017'
Samples$Year='2018'
Samples=rbind(Samples,PupilsSamples)

Fixations$Ostension[Fixations$Ostension=='+o']='Ostensive'
Fixations$Ostension[Fixations$Ostension=='-o']='Private'
Samples$Ostension[Samples$Ostension=='+o']='Ostensive'
Samples$Ostension[Samples$Ostension=='-o']='Private'

Fixations$Directionality[Fixations$Directionality=='div']='Towards3rd'
Fixations$Directionality[Fixations$Directionality=='dir']='TowardsYou'
Samples$Directionality[Samples$Directionality=='div']='Towards3rd'
Samples$Directionality[Samples$Directionality=='dir']='TowardsYou'

Fixations$ActorGender[Fixations$ActorGender=='f']='Female'
Fixations$ActorGender[Fixations$ActorGender=='m']='Male'
Samples$ActorGender[Samples$ActorGender=='f']='Female'
Samples$ActorGender[Samples$ActorGender=='m']='Male'

Fixations$SearchOrder[Fixations$SearchOrder=='First']='1'
Fixations$SearchOrder[Fixations$SearchOrder=='Second']='2'
Saccades$SearchOrder[Saccades$SearchOrder=='First']='1'
Saccades$SearchOrder[Saccades$SearchOrder=='Second']='2'

write_csv(Saccades,"SaccadesV3.csv")
write_csv(Fixations,"FixationsV3.csv")
write_csv(Samples,"SamplesV3.csv")
```

## Now plotting the data (only 2018 data here, use V3 if you want to visualize them all)

```{r plots}

SamplesV2=read_csv('SamplesV2.csv')
FixationsV2=read_csv('FixationsV2.csv')
SaccadesV2=read_csv('SaccadesV2.csv')


FixationsV2$PositionYClean=1081-FixationsV2$PositionY


#SANITY CHECKS

#check for outliers
ggplot(FixationsV2, aes(ParticipantID, Duration)) +
  geom_boxplot() +
  facet_grid(.~Task)

#check distribution of fixations
# FixationsV2$Duration=as.numeric(FixationsV2$Duration)
hist(FixationsV2$Duration[FixationsV2$Task=='VisualSearch']) #lognormal-ish
hist(FixationsV2$Duration[FixationsV2$Task=='SocialEngagement']) #lognormal-ish

ggplot(FixationsV2, aes(Duration,color=Task)) + geom_density()

#check distribution of saccades
hist(SaccadesV2$Amplitude[SaccadesV2$Task=='VisualSearch']) #lognormal-ish
hist(SaccadesV2$Amplitude[SaccadesV2$Task=='SocialEngagement']) #lognormal-ish
hist(SaccadesV2$MeanVelocity[SaccadesV2$Task=='VisualSearch']) #lognormal-ish
hist(SaccadesV2$MeanVelocity[SaccadesV2$Task=='SocialEngagement']) #lognormal-ish
hist(SaccadesV2$PeakVelocity[SaccadesV2$Task=='VisualSearch']) #lognormal-ish
hist(SaccadesV2$PeakVelocity[SaccadesV2$Task=='SocialEngagement']) #lognormal-ish

#check distribution of pupil size 
hist(SamplesV2$PupilSize[SamplesV2$Task=='SocialEngagement']) #multimodal?
ggplot(subset(SamplesV2,Task=='SocialEngagement'),aes(PupilSize,color=ParticipantID))+geom_density() # Not really
hist(SamplesV2$PupilSize[SamplesV2$Task=='VisualSearch']) #bimodal?
ggplot(subset(SamplesV2,Task=='VisualSearch'),aes(PupilSize,color=ParticipantID))+geom_density() # Not really!

#RUN SOME ANALYSES AND VISUALIZATIONS

#visualize VELOCITY over trial time for ONE participant/trial
ggplot(data=subset(SamplesV2,ParticipantID=='1_1_f1' & Trial==1), aes(x=TrialTime, y=abs(MeanVelocityX))) +
  geom_line()+
  geom_point() # + facet_grid(.~ParticipantGender)

#visualize ACCELERATION over trial time for ONE participant/trial
ggplot(data=subset(SamplesV2,ParticipantID=='1_1_f1' & Trial==1), aes(x=TrialTime, y=abs(MeanAccellerationX))) +
  geom_line()+
  geom_point() # + facet_grid(.~ParticipantGender)

#visualize SACCADE AMPLITUDE (difference between VisualSeach and SocialEngagement in terms of saccades)
ggplot(data=SaccadesV2, aes(x=Saccade,y=Amplitude,group=ParticipantID)) +
  geom_line(aes(color=ParticipantID))+
  geom_point(aes(color=ParticipantID))+
  facet_grid(.~Task)

#visualize FIXATION DURATION (difference between VisualSeach and SocialEngagement in terms of fixations)
ggplot(data=FixationsV2, aes(x=Fixation,y=Duration,group=ParticipantID)) +
  geom_line(aes(color=ParticipantID))+
  geom_point(aes(color=ParticipantID))+
  facet_grid(.~Task)

#visualize FIXATION DURATION BY GENDER
ggplot(data=subset(FixationsV2, Trial==1), aes(x=Fixation,y=Duration)) +
  geom_point()+
  geom_line()+
  geom_smooth() +
  facet_grid(.~ParticipantGender)

library(jpeg)
library(grid)

# HEATMAPS (https://gist.github.com/kalebr/6832380)
jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
img <- jpeg::readJPEG('EyeTrackingScripts/foraging/ng090ws.jpg')
g <- grid::rasterGrob(img, interpolate=TRUE)

#density
ggplot(subset(FixationsV2, Task=='VisualSearch' & ParticipantID=='6_3_m2' & Trial==1), aes(x = PositionX, y = PositionYClean)) +
  xlim(0,1920) +
  ylim(0, 1080) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=0, ymax=1080) + #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  stat_density2d(geom="raster", aes(fill=..density.., alpha=sqrt(sqrt(..density..))), contour=FALSE, n=1000) + 
  scale_alpha(range = c(0.1, 0.6)) + scale_fill_gradientn(colours = jet.colors(10), trans="sqrt")

x=subset(FixationsV2, ParticipantID=='2_2_f2' & Trial==4)
x= x[order(x$Fixation),]

img <- readJPEG("EyeTrackingScripts/foraging/ng064ws.jpg")
g <- rasterGrob(img, interpolate=TRUE)

#SCANPATHS
ggplot(x, aes(x=PositionX, y=PositionYClean, label=Fixation)) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-0, ymax=1080) + #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  geom_point(size = x$Duration[x$ParticipantID=='2_2_f2' & x$Trial==4]/100, alpha = .7, color='red') +
  geom_path(size = 1, alpha = 0.3,color='red') +
  geom_text(aes(label = Fixation, size = 5)) +
  labs(x="postion x", y="position y")


x=subset(FixationsV2, ParticipantID=='2_2_f2' & Trial==9)
x= x[order(x$Fixation),]

img <- readJPEG("EyeTrackingScripts/foraging/ng151ws.jpg")
g <- rasterGrob(img, interpolate=TRUE)

#SCANPATHS
ggplot(x, aes(x=PositionX, y=PositionYClean, label=Fixation)) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-0, ymax=1080) + #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  geom_point(size = x$Duration[x$ParticipantID=='2_2_f2' & x$Trial==9]/100, alpha = .7, color='red') +
  geom_path(size = 1, alpha = 0.3,color='red') +
  geom_text(aes(label = Fixation, size = 5)) +
  labs(x="postion x", y="position y")

#GROWTH PLOTS PUPIL SIZE
ggplot(subset(SamplesV2, Task=='SocialEngagement'),   #evt. FixationsV2,StartTime
       aes(TrialTime, PupilSize, colour=ParticipantGender), na.rm=T) +
       geom_smooth() +
       xlim(0,5371)+
       facet_grid(.~ActorGender) #Task #Ostension #ActorGender #Directionality

ggplot(subset(SamplesV2, Task=='SocialEngagement'),   #evt. FixationsV2,StartTime
       aes(TrialTime, PupilSize, colour=Directionality), na.rm=T) +
       geom_smooth() +
       xlim(0,5371)+
       facet_grid(.~Ostension) #Task #Ostension #ActorGender #Directionality
```

```{r}


# HEATMAPS (https://gist.github.com/kalebr/6832380)
library(jpeg)
library(grid)

jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
img <- readJPEG("EyeTrackingScripts/foraging/ng064ws.jpg")
g <- rasterGrob(img, interpolate=TRUE) #to load the image and get it pop up behind the plot

#density
ggplot(subset(FixationsV2, Task=='VisualSearch' & ParticipantID=='2_2_f2' & Trial==4), aes(x = PositionX, y = PositionYClean)) +
  xlim(0,1920) + #the screen resolution
  ylim(0, 1080) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=0, ymax=1080) + 
  #specify how big the picture should be #not same resolution, the picture can't be fitted using both
  #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  stat_density2d(geom="raster", aes(fill=..density.., alpha=sqrt(sqrt(..density..))), contour=FALSE, n=1000) + 
  scale_alpha(range = c(0.1, 0.6)) + scale_fill_gradientn(colours = jet.colors(10), trans="sqrt") +
  labs(x="postion x", y="position y")

img <- readJPEG("EyeTrackingScripts/foraging/ng151ws.jpg")
g <- rasterGrob(img, interpolate=TRUE) #to load the image and get it pop up behind the plot

#density
ggplot(subset(FixationsV2, Task=='VisualSearch' & ParticipantID=='2_2_f2' & Trial==9), aes(x = PositionX, y = PositionYClean)) +
  xlim(0,1920) + #the screen resolution
  ylim(0, 1080) +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=0, ymax=1080) + 
  #specify how big the picture should be #not same resolution, the picture can't be fitted using both
  #xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  stat_density2d(geom="raster", aes(fill=..density.., alpha=sqrt(sqrt(..density..))), contour=FALSE, n=1000) + 
  scale_alpha(range = c(0.1, 0.6)) + scale_fill_gradientn(colours = jet.colors(10), trans="sqrt") +
  labs(x="postion x", y="position y")



```


1. Visual Search

```{r}

Fixations=read_csv('FixationsV2.csv')
Saccades=read_csv('SaccadesV2.csv')
Samples=read_csv('SamplesV2.csv')

# Visual Search
#exclude social
Fixations_vis= subset(Fixations, c(Task== "VisualSearch"))

# Create extra ID
Fixations_vis$extra_id= as.numeric(as.factor(as.character(Fixations_vis$ParticipantID)))

# LogTransform Duration
Fixations_vis$Duration_log=log(Fixations_vis$Duration)
library(lmerTest)
#models for fixation duration
normal = lmer(Duration_log ~ SearchType + Trial + (1 + SearchType|extra_id) ,data= Fixations_vis)
summary(normal)

inter= lmer(Duration_log ~ SearchType * Trial + (1 + SearchType|extra_id),data= Fixations_vis)
summary(inter)

simple= lmer(Duration_log ~ SearchType + (1 + SearchType|extra_id),data= Fixations_vis)
summary(simple)

simpler= lmer(Duration_log ~ SearchType + (1|extra_id), data= Fixations_vis)
summary(simpler)

models_fix=list(normal, inter, simple, simpler)
```

Crossvalidation

```{r}

#VISUAL SEARCH

Fixations_vis$ID=as.numeric(as.factor(as.character((Fixations_vis$ParticipantID))))
             
#building function for cross validation for fixations
crossvalidation_fix  = function (model) {
  "%not in%" <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))
  Folds = caret::createFolds(unique(Fixations_vis$ID),3)
  result={}
  for (i in Folds) {
    #subset of Fixations_visa except for the 1 fold
    train_d = subset(Fixations_vis, ID %not in% i)
    #subset including only 1 fold
    test_d = subset(Fixations_vis, ID %in% i)
    #fit train subset into specified model
    model_val = update(model,data=train_d)
    #rmse of train data
    pred1 = predict(model_val, train_d,allow.new.levels=T)
    rmse_train=Metrics::rmse(train_d$Duration_log[!is.na(pred1)],pred1[!is.na(pred1)])
    
    #rmse of test data
    pred2 = predict(model_val, test_d,allow.new.levels=T)
    rmse_test=Metrics::rmse(test_d$Duration_log[!is.na(pred2)],pred2[!is.na(pred2)])
    
    res = cbind(rmse_train,rmse_test)
    result=rbind(result,res)
  }
  
  final_res=list(mean_train=mean(result[1]),mean_test=mean(result[2]))
  return(final_res)
}

```

```{r crossvalidation}

hola=data.frame(mean_test=numeric(), mean_train=numeric())

for (model in models_fix) {
  final=data.frame()
  n=0
  while(n < 20) {
  #cross validate function
  cross=crossvalidation_fix(model)
  final=rbind(final, cross)
  n=n+1
  }
  hola[nrow(hola)+1,1]=mean(final[,1])
  hola[nrow(hola),2]=mean(final[,2])
} 

write.csv(final, file= "models_fix.csv", row.names = F)
fix_models=read.csv("models_fix.csv")

```


```{r amplitude models}

# Saccade amplitude
Saccades_vis= subset(Saccades, c(Task== "VisualSearch"))
Saccades_vis$extra_id= as.numeric(as.factor(as.character(Saccades_vis$ParticipantID)))
# LogTransform Duration
Saccades_vis$Amplitude_log=log(Saccades_vis$Amplitude)
Saccades_vis$SearchType=as.factor(Saccades_vis$SearchType)
library(lmerTest)
normal_s = lmer(Amplitude_log ~ SearchType + Trial + (1 + SearchType|extra_id) ,data= Saccades_vis)
summary(normal_s)

inter_s= lmer(Amplitude_log ~ SearchType * Trial + (1 + SearchType|extra_id),data= Saccades_vis)
summary(inter_s)

simple_s= lmer(Amplitude_log ~ SearchType + (1 + SearchType|extra_id),data= Saccades_vis)
summary(simple_s)

simpler_s= lmer(Amplitude_log ~ SearchType + (1|extra_id), data= Saccades_vis)
summary(simpler_s)

models_s=list(normal_s, inter_s, simple_s, simpler_s)

levels(Saccades_vis$SearchType)
```

```{r crossval function for saccades}
             
#building function for cross validation for saccades
crossvalidation_sac  = function (model) {
  "%not in%" <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))
  Folds = caret::createFolds(unique(Saccades_vis$extra_id),3)
  result={}
  for (i in Folds) {
    #subset of Fixations_visa except for the 1 fold
    train_d = subset(Saccades_vis, extra_id %not in% i)
    train_d=subset(train_d, !is.na(train_d$Amplitude)) # without this, it gives back NA-s
    #subset including only 1 fold
    test_d = subset(Saccades_vis, extra_id %in% i)
    test_d=subset(train_d, !is.na(test_d$Amplitude)) # without this, it gives back NA-s
    #fit train subset into specified model
    model_val = update(model, data=train_d)
    #rmse of train data
    pred1 = predict(model_val, train_d,allow.new.levels=T)
    rmse_train=Metrics::rmse(train_d$Amplitude_log[!is.na(pred1)],pred1[!is.na(pred1)])
    
    #rmse of test data
    pred2 = predict(model_val, test_d,allow.new.levels=T)
    rmse_test=Metrics::rmse(test_d$Amplitude_log[!is.na(pred2)],pred2[!is.na(pred2)])
    
    res = cbind(rmse_train,rmse_test)
    result=rbind(result,res)
    
  }
  
  final_res=list(mean_train=mean(result[1]),mean_test=mean(result[2]))
  return(final_res)
}
```

```{r crossval}

uj=data.frame(mean_test=numeric(), mean_train=numeric())
for (model in models_s) {
  final_sac=data.frame()
  n=0 
  while(n < 20) {
    #cross validate function
    cross=crossvalidation_sac(model) # próbáld meg beleírni a nevet
    final_sac=rbind(final_sac, cross)
    n=n+1
  }
  uj[nrow(uj)+1,1]=mean(final_sac[,1])
  uj[nrow(uj),2]=mean(final_sac[,2])
  
}

write.csv(final_sac, file= "models_sac.csv", row.names = F)
sac_models=read.csv("models_sac.csv")

# simple_s a legjobb
```

2. Social engagement

```{r social engagement models}
# with samples
library(dplyr)
Samples_soc=Samples %>% subset(Task== "SocialEngagement") %>% 
  subset(!is.na(Fixation)) %>% subset(!is.na(PupilSize))

# Samples_soc$ID=as.numeric(as.factor(as.character((Samples_soc$ParticipantID)))) 
Samples_soc$extra_id= as.numeric(as.factor(as.character(Samples_soc$ParticipantID)))
Samples_soc$TrialTime=Samples_soc$
Samples_soc$scaled_TrialTime=scale(Samples_soc$TrialTime)
library(lmerTest)
# models
simpler_pup = lmer(PupilSize ~ Ostension + Directionality + scaled_TrialTime + (1 | extra_id) ,data= Samples_soc, control = lmerControl(calc.derivs = FALSE))
summary(simpler_pup)

simple_pup = lmer(PupilSize ~ Ostension + Directionality + scaled_TrialTime + (1 + scaled_TrialTime + Ostension + Directionality | extra_id) ,data= Samples_soc, control = lmerControl(calc.derivs = FALSE))
summary(simple_pup)

one = lmer(PupilSize ~ Ostension * Directionality * scaled_TrialTime + (1 + scaled_TrialTime + Ostension + Directionality | extra_id) ,data= Samples_soc, control = lmerControl(calc.derivs = FALSE))
summary(one)

two = lmer(PupilSize ~ Ostension * Directionality * (scaled_TrialTime + I(scaled_TrialTime^2)) + (1 + scaled_TrialTime + Ostension + Directionality | extra_id) ,data= Samples_soc, control = lmerControl(calc.derivs = FALSE))
summary(two)

three = lmer(PupilSize ~ Ostension * Directionality * (scaled_TrialTime + I(scaled_TrialTime^2) + I(scaled_TrialTime^3)) + (1 + scaled_TrialTime + Ostension + Directionality | extra_id) ,data= Samples_soc, control = lmerControl(calc.derivs = FALSE))
summary(three)

models_pup=list(simpler_pup, simple_pup, one, two, three)

```

```{r function for crossval for pupilsize}
#building function for cross validation for pupilsize
crossvalidation_pup  = function (model) {
  "%not in%" <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))
  Folds = caret::createFolds(unique(Samples_soc$extra_id),3)
  result={}
  for (i in Folds) {

    train_d = subset(Samples_soc, extra_id %not in% i)
    train_d=subset(train_d, !is.na(train_d$PupilSize)) # without this, it gives back NA-s
    #subset including only 1 fold
    test_d = subset(Samples_soc, extra_id %in% i)
    test_d=subset(train_d, !is.na(test_d$PupilSize)) # without this, it gives back NA-s
    #fit train subset into specified model
    model_val = update(model, data=train_d)
    #rmse of train data
    pred1 = predict(model_val, train_d,allow.new.levels=T)
    rmse_train=Metrics::rmse(train_d$PupilSize[!is.na(pred1)],pred1[!is.na(pred1)])
    
    #rmse of test data
    pred2 = predict(model_val, test_d,allow.new.levels=T)
    rmse_test=Metrics::rmse(test_d$PupilSize[!is.na(pred2)],pred2[!is.na(pred2)])
    
    res = cbind(rmse_train,rmse_test)
    result=rbind(result,res)
    
  }
  
  final_res=list(mean_train=mean(result[1]),mean_test=mean(result[2]))
  return(final_res)
}
```

```{r crossvalidation}

new=data.frame(mean_test=numeric(), mean_train=numeric())
for (model in models_pup) {
  final_pup=data.frame()
  n=0 
  while(n < 1) {
    #cross validate function
    cross=crossvalidation_pup(model)
    final_pup=rbind(final_pup, cross)
    n=n+1
  }
  new[nrow(new)+1,1]=mean(final_pup[,1])
  new[nrow(new),2]=mean(final_pup[,2])
  
}

write.csv(final_pup, file= "models_pup.csv", row.names = F)
pup_models=read.csv("models_pup.csv")
```



